---
title: "World Bank Capstone Project Data Viz"
author: "Qi Xue, Neha Dagaonkar, James Hawkins"
date: "11/2/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(haven)
library(here)
library(scales)
library(reshape)
library(reshape2)
library(data.table)
library(formattable)
library(kableExtra)
library(sjlabelled)
library(ggalt)
```

### Figure 1. Evolution in number of countries/territories and social protection measures
```{r figure 1}
# load output data
figure1 <- read_dta(here("output/", "figure_1.dta")) %>% 
  drop_na()

# plot the graph
figure1 %>% 
   ggplot() + 
   geom_col(aes(x = cum_month, y = cum_num_country), fill = "darkgoldenrod2") +
     geom_text(aes(x = cum_month, y = cum_num_country,
                   label = cum_num_country), 
                   vjust = - 1, size = 2.5) +
   geom_line(aes(x = cum_month, y = cum_num_measure/4), color = "darkblue") +
     # geom_text(aes(x = cum_month, y = cum_num_measure/4,
     #               label = cum_num_measure),
     #               vjust = - 1, size = 2.5) +
     scale_y_continuous(sec.axis = sec_axis(~.*4, name = "Number of new social protection measures")) +
     # scale_x_continuous(labels = c("May2020", "Jun2020", "Jul2020", "Aug2020", "Sep2020", "Oct2020", "Nov2020", "Dec2020",
     #                               "Jan2021", "Feb2021", "Mar2021", "Apr2021", "May2021", "Jun2021", "Jul2021", "Aug2021",
     #                               "Sep2021", "Oct2021", "Nov2021", "Dec2021")) +
   theme_minimal() +
   labs(title = "Figure 1. Evolution in number of countries/territories and social protection measures",
         x = "Number of months since Mar 2020",
         y = "Number of countries/territories (bars)") +
    theme(plot.title = element_text(face = "bold", size = 11, vjust = 2), 
         axis.title.x = element_text(size = 9),
         axis.title.y = element_text(size = 9))

ggsave(here("output/data_viz", "figure_1.png"))
```

### Figure 2: Composition of social protection response by components in regions and country income groups
```{r figure 2, fig.width = 7, fig.height = 4}
library(patchwork)

# load output data1
figure2_1 <- read_dta(here("output/", "figure_2_part1.dta")) %>% 
  drop_na()

# melt data
melt_figure2_1 <- reshape2::melt(figure2_1, id = c("region"))

# plot the graph
P1 <- melt_figure2_1 %>%
  ggplot(aes(x = region, y = value, fill = variable)) +
  geom_col(position = "fill", width = 0.6) +
  scale_fill_manual(values = c("lightgrey", "navy", "darkgoldenrod2")) +
  geom_text(aes(label = round(value, 1)), 
            color = "white",
            size = 2.5,
            position = position_fill(vjust = 0.5)) +
  scale_y_continuous(breaks = c(0.00, 0.25, 0.50, 0.75, 1.00),
                     labels = c("0", "25%", "50%", "75%", "100%")) +
  theme_minimal() +
  labs(title = "Figure 2: Composition of social protection response by components in regions and country income groups",
       x = "", y = "") +
  theme(plot.title = element_text(face = "bold", size = 10, vjust = 1)) +
  theme(legend.position = "none")

# load output data2
figure2_2 <- read_dta(here("output/", "figure_2_part2.dta")) %>% 
  drop_na()

# melt data
melt_figure2_2 <- reshape2::melt(figure2_2, id = c("income_group"))

# plot the graph
P2 <- melt_figure2_2 %>%
  ggplot(aes(x = income_group, y = value, fill = variable)) +
  geom_col(position = "fill", width = 0.7) +
  geom_text(aes(label = round(value, 1)), 
            color = "white",
            size = 2.5,
            position = position_fill(vjust = 0.5)) +
  theme_minimal() + 
  labs(x = "", y = "") +
  theme(plot.title = element_text(face = "bold", size = 11, vjust = 1, hjust = -0.5)) +
  theme(axis.text.y = element_blank(),
        legend.position = "right",
        legend.title = element_blank()) +
  scale_fill_manual(breaks = c("pct_assistance", "pct_insurance", "pct_labor"),
                    labels = c("Social Assistance\n", "Social Insurance\n", "Labor Market"),
                    values = c("lightgrey", "navy", "darkgoldenrod2"))

P1 + P2 + plot_layout(ncol = 2, width = c(2, 1))

ggsave(here("output/data_viz", "Figure_2.png"))
```
### Table 1. Social assistance measures
```{r table1, fig.width = 3}

# load output data1
table1 <- read_dta(here("output/", "table_1.dta"))

# get encode labels
labels <- get_labels(table1$sp_category_code)
table1$sp_category_code <- map_chr(.x = table1$sp_category_code, .f = ~labels[.x])

# formalize table
table1 <-table1 %>% 
  mutate(`Social Protection Categories` = str_remove(sp_category_code,"[1].[1-9]. "),
         `Number of Measures` = N_of_measures,
         `Number of Countries` = N_of_countries) %>% 
  select(-sp_category_code, -N_of_measures, -N_of_countries) 

# present the table
formattable(table1,
            align = c("l","c","c"), list(
  `Social Protection Categories` = formatter("span", 
                               style = ~ style(color = "grey", font.weight = "bold")), 
  `Number of Measures` = color_tile("#DeF7E9", "#82d58c"),
  `Number of Countries` = color_tile("lightblue", "lightblue3")))

# alternatively, using the length of bar to indicate numbers
# formattable(table3,
#             align = c("l","c","c"), list(
#   `Social Protection Categories` = formatter("span", 
#                                style = ~ style(color = "grey", font.weight = "bold")), 
#    area(col = c(`Number of Measures`, `Number of Countries`)) ~ normalize_bar("pink", 0.2)))
```

### Table 2. Social insurance programs
```{r table2}

# load output data
table2 <- read_dta(here("output/", "table_2.dta")) %>% 
  drop_na() 

# get encode labels
labels <- get_labels(table2$sp_category_code)
table2$sp_category_code <- map_chr(.x = table2$sp_category_code, .f = ~labels[.x])

# formalize table
table2 <-table2 %>% 
  mutate(`Social Protection Categories` = str_remove(sp_category_code,"[1-9].[1-9]. "),
         `Number of Measures` = N_of_measures,
         `Number of Countries` = N_of_countries) %>% 
  select(-sp_category_code, -N_of_measures, -N_of_countries) 

# present the table
formattable(table2,
            align = c("l","c","c"), list(
  `Social Protection Categories` = formatter("span", 
                               style = ~ style(color = "grey", font.weight = "bold")), 
  `Number of Measures` = color_tile("#DeF7E9", "#82d58c"),
  `Number of Countries` = color_tile("lightblue", "lightblue3")))
```

### Table 3. Active Labor market and labor regulations
```{r table3}

# load output data
table3 <- read_dta(here("output/", "table_3.dta")) %>% 
  drop_na() 

# get encode labels
labels <- get_labels(table3$sp_category_code)
table3$sp_category_code <- map_chr(.x = table3$sp_category_code, .f = ~labels[.x])

# formalize table
table3 <- table3 %>% 
  mutate(`Social Protection Categories` = str_remove(sp_category_code,"[3].[1-9]. "),
         `Number of Measures` = N_of_measures,
         `Number of Countries` = N_of_countries) %>% 
  select(-sp_category_code, -N_of_measures, -N_of_countries) 

# present the table
formattable(table3,
            align = c("l","c","c"), list(
  `Social Protection Categories` = formatter("span", 
                               style = ~ style(color = "grey", font.weight = "bold")), 
  `Number of Measures` = color_tile("#DeF7E9", "#82d58c"),
  `Number of Countries` = color_tile("lightblue", "lightblue3")))
```

### Table 4. State in social assistance implementation
```{r table4}

# load output data
table4 <- read_dta(here("output/","table_4.dta")) %>% 
  select(c(`sp_category_code`, `Planned`, `Ongoing`, `Ended`)) %>%
  drop_na() 

# get encoded labels
labels <- get_labels(table4$sp_category_code)
table4$sp_category_code <- map_chr(.x = table4$sp_category_code, .f = ~labels[.x])

# formalize table
table4 <-table4 %>% 
  mutate(`Social Protection Categories` = str_remove(sp_category_code,"[1-9].[1-9]. "),
         `Programs Planned` = Planned,
         `Programs Ongoing` = Ongoing,
         `Programs Ended` = Ended) %>% 
  select(-sp_category_code, -Planned, -Ongoing, -Ended) 

# present the table
formattable(table4,
            align = c("l","c","c"), list(
  sp_category_code = formatter("span", 
                               style = ~ style(color = "grey", font.weight = "bold")), 
  `Programs Planned` = color_tile("#DeF7E9", "#82d58c"),
  `Programs Ongoing` = color_tile("lightblue", "lightblue3"),
  `Programs Ended` = color_tile("pink", "pink3")))
  
#labs(title = paste("Table 4. State in social assistance implementation"))
```

### Figure 3. Generosity of cash transfer size relative to pre-COVID level
```{r figure 3, fig.width = 9, fig.height = 5.5}

# load output data
figure3 <- read_dta(here("output/", "figure_3.dta")) %>% 
  drop_na()

# combine label
figure3$combined <- 
  paste(figure3$country_name, figure3$program_name, sep = " (")

figure3$combined <-
  paste(figure3$combined, ")", sep = "")

figure3 <- figure3 %>% 
  mutate(combined = str_remove(combined, " \\(\\)"))

# plot the graph
figure3 %>% 
  mutate(combined = forcats::fct_reorder(.x = sa_transfer,
                                         .f = combined)) %>% 
  ggplot() +
  geom_col(aes(x = combined, y = sa_transfer),
           fill = "darkgoldenrod2") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Figure 3. Generosity of cash transfer size relative to pre-COVID level",
       y = "% increase in transfer size from pre-covid level",
       x = "") +
  # geom_text(aes(x = combined, y = sa_transfer,
  #               label = round(sa_transfer*100, 2)),
  #               hjust = -0.5, size = 3) +
  scale_y_continuous(breaks = 0:4,
                     labels = scales::percent) +
  theme(plot.title = element_text(face = "bold", size = 11, vjust = 1),
        axis.title.x = element_text(size = 10))

ggsave(here("output/data_viz", "figure_3.png"))
```

### Table 5. Adaption of cash transfer program
```{r table 5}

# load output data
table5 <- read_dta(here("output/", "table_5.dta")) %>% 
  drop_na() 
```

### Figure 4. Duration in cash transfer programs
```{r figire 4}

# load output data
figure4 <- read_dta(here("output/", "figure_4.dta")) %>% 
  drop_na() 

# create month column
figure4 <- figure4 %>% 
  mutate(month = ifelse(duration <= 1, "month", "months"))

# paste months
figure4 <- figure4 %>% 
  mutate(duration_month = paste(figure4$duration, figure4$month))

# plot the graph
figure4 %>% 
  ggplot() +
  geom_col(aes(x = factor(duration_month, 
                          levels = c("0 month", "1 month", 
                                     "2 months", "3 months",
                                     "4 months", "5 months", 
                                     "6 months", "7 months",
                                     "8 months", "9 months")), 
              y = num_projects), 
              fill = "lightblue3", 
              width = 0.6) +
  coord_flip() +
  theme_light() +
  geom_vline(xintercept = 3.84, color = "red",
             show.legend = "Average") + # name is not showing up
  labs(title = "Figure 4. Duration in cash transfer programs",
       y = "Number of cash transfer schemes",
       x = "") +
  scale_y_continuous(breaks = 0: 20) +
  theme(plot.title = element_text(face = "bold", size = 11, vjust = 1.5, hjust = 0.5),
         axis.title.x = element_text(size = 10))

ggsave(here("output/data_viz", "figure_4.png"))
```

### Figure 5. Planned versus actual coverage rates for top 50 programs
```{r figure 5, fig.width = 9.5, fig.height = 5.5}

# load data
figure5 <- read_dta(here("output/", "figure_5.dta")) %>%
    mutate(name = str_remove(program_name,"[0-9][0-9]. ")) %>% 
    mutate(name = str_remove(name,"[0-9]. "))# remove number in program name

# paste program name to country name
figure5$combined <- 
  paste(figure5$country_name, figure5$name, sep = " (")

figure5$combined <-
  paste(figure5$combined, ")", sep = "")

# plot the data
figure5 %>% 
  mutate(combined = forcats::fct_reorder(.x = sa_app,
                                         .f = combined)) %>% 
  ggplot() +
  geom_segment(aes(y = combined, yend = combined, 
                   x = 0, xend = .5), 
               color="#b2b2b2", size = 0.1) +
  geom_dumbbell(aes(y = combined, x = sa_app, xend = sa_ppp),
                size = 1, color = "#b2b2b2", alpha = 0.8,
                size_x = 2.5, size_xend = 2.5, 
                colour_x = "navy", colour_xend = "darkgoldenrod2", 
                show.legend = TRUE) +
  theme_bw() +
  scale_color_manual(name = "", values = c("navy", "darkgoldenrod2")) +
  #theme_minimal() +
  labs(title = "Figure 5. Planned versus actual coverage rates for top 50 programs",
       x = "", y = "") +
  theme(plot.title = element_text(face = "bold", size = 11, vjust = 2), 
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9))

ggsave(here("output/data_viz", "figure_5.png"))
```

### Table 6. Number of social assistance and cash transfer beneficiaries (planned and actual)
```{r table 6}

# load output data
table6 <- read_dta(here("output/", "table_6.dta"))

# rename var name
table6 <- table6 %>% 
   mutate(`Planned number of individuals` = prettyNum(ben_totals0, big.mark = ','),
          `Planned number of inviduals as % of the World's population` = round(ben_percent0, 2),
          `Actual number of individuals` = prettyNum(ben_totals1,big.mark = ','),
          `Actual number of inviduals as % of the World's population` = round(ben_percent1, 2)) %>% 
  select(-ben_totals0, -ben_totals1, -ben_percent0, -ben_percent1) 

# get encode labels 
labels <- get_labels(table6$y)
table6$y <- map_chr(.x = table6$y, .f = ~labels[.x])

# present the table
formattable(table6,
            align = c("l","c","c"), list(
   y = formatter("span", style = ~ style(color = "grey", font.weight = "bold")),
   `Planned number of individuals` = color_tile("#DeF7E9", "#82d58c"),
   `Actual number of individuals` = color_tile("lightblue", "lightblue3")))
```

### Figure 6. Top 10 cash transfer program by actual coverage (million people)
```{r figure 6, fig.width = 8, fig.height = 3}

# load output data
figure6 <- read_dta(here("output/", "figure_6.dta")) %>% 
   drop_na() 

# combine label
figure6$combined <- 
  paste(figure6$country_name, figure6$program_name, sep = " (")

figure6$combined <-
  paste(figure6$combined, ")", sep = "")

figure6 <- figure6 %>% 
  mutate(combined = str_remove(combined, " \\(\\)"),
        combined = str_remove(combined,"[1][0-9]. "))

# plot the graph
figure6 %>% 
  mutate(ben_actual_m = ben_actual/1000000) %>% 
  mutate(combined = forcats::fct_reorder(.x = ben_actual_m, .f = combined)) %>% 
  ggplot() +
  geom_col(aes(x = ben_actual_m, y = combined),
           fill = "darkgoldenrod2",
           width = 0.4) +
  theme_minimal() +
  labs(title = "Figure 6. Top 10 cash transfer program by actual coverage (million people)",
       y = "",
       x = "") +
   theme(plot.title = element_text(face = "bold", size = 11, vjust = 1, hjust = 1))

ggsave(here("output/data_viz", "figure_6.png"))
```

### Figure 7. Top 10 cash transfer programs by actual coverage (% population)
```{r figure 7, fig.width = 7, fig.height = 3}

# load output data
figure7 <- read_dta(here("output/", "figure_7.dta")) %>% 
  drop_na() 

# get encode labels
# labels7 <- get_labels(figure7$sa_policy_adap2)
# figure7$sa_policy_adap2 <- map_chr(.x = figure7$sa_policy_adap2, .f = ~labels[.x])

# remove numbers from label
# figure7 <- figure7 %>% 
  # mutate(sa_policy_adap2 = str_remove(sa_policy_adap2,"[1-9].[1-9]. "))

# combine label
figure7$combined <- 
  paste(figure7$country_name, figure7$sa_policy_adap, sep = " (")

figure7$combined <-
  paste(figure7$combined, ")", sep = "")

# plot the graph
figure7 %>% 
  mutate(combined = forcats::fct_reorder(.x = sa_app,
                                         .f = combined)) %>% 
  ggplot() +
  geom_col(aes(x = combined, y = sa_app),
           fill = "darkgoldenrod2",
           width = 0.4) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Figure 7. Top 10 cash transfer programs by actual coverage (% population)",
       y = "% of population covered by program",
       x = "") +
  theme(plot.title = element_text(face = "bold", size = 11, vjust = 1, hjust = 1.2),
        axis.title.x = element_text(size = 10))

ggsave(here("output/data_viz", "figure_7.png"))
```

### Figure 8. Scale up rates of cash transfer relative to pre-COVID level
```{r figure 8}

# load output data
figure8 <- read_dta(here("output/", "figure_8.dta")) %>% 
  drop_na() 

# plot the graph
figure8 %>% 
  mutate(country_name = forcats::fct_reorder(.f = country_name,
                                             .x = rate_scaleup)) %>% 
  ggplot() +
  geom_col(aes(x = country_name, y = rate_scaleup),
           fill = "lightblue3",
           width = 0.5) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Figure 8. Scale up rates of cash transfer relative to pre-COVID level",
       y = "% increase in transfer size from pre-covid level",
       x = "") +
  theme(plot.title = element_text(face = "bold", size = 11, vjust = 1, hjust = 0),
        axis.title.x = element_text(size = 10)) 

ggsave(here("output/data_viz", "figure_8.png"))
```
### Table 7. Spending by social protection component and region group (USD)
```{r table7}

options(scipen = 99)

# load output data
table7 <- read_dta(here("output/","table_7.dta")) %>% 
    select(c(`inc_grp_code`, `exp_plan_num_usd1`, `exp_plan_num_usd2`,`exp_plan_num_usd3`, `total_expenditure`, `expenditure_per_capita`)) %>%
  drop_na() %>% 
  mutate(total_expenditure = round(total_expenditure, 2))

# get encoded labels
labels <- get_labels(table7$inc_grp_code)
table7$inc_grp_code <- map_chr(.x = table7$inc_grp_code, .f = ~labels[.x])

# formalize table
table7 <-table7 %>% 
  mutate(`Countries by Income Group` = inc_grp_code,
         `Social Assistance` = exp_plan_num_usd1,
         `Social Insurance` = exp_plan_num_usd2,
         `Labor Markets` = exp_plan_num_usd3,
         `Total Spending (USD)` = round(total_expenditure, 2),
         `Spending pc (average $/capita)` = expenditure_per_capita) %>% 
  select(-inc_grp_code, -exp_plan_num_usd1, -exp_plan_num_usd2, -exp_plan_num_usd3, -total_expenditure, -expenditure_per_capita) 

# present the table
formattable(table7,
            align = c("l","c","c","c","c","c"), list(
  inc_grp_code = formatter("span", 
                               style = ~ style(color = "grey", font.weight = "bold")), 
  `Social Assistance` = color_tile("#DeF7E9", "#82d58c"),
  `Social Insurance` = color_tile("#DeF7E9", "#82d58c"),
  `Labor Markets` = color_tile("#DeF7E9", "#82d58c"), 
  `Total Spending (USD)` = color_tile("lightblue", "lightblue3"),
  `Spending pc (average $/capita)` = color_tile("lightblue", "lightblue3")))
  
# labs(title = paste("Table 7. Spending by social protection component - country income group (USD)"))
```

### Pie chart
```{r pie chart in excel, eval = FALSE, include = FALSE}

# load output data
figure9.1 <- read_dta(here("output/", "figure_9.dta")) %>% 
  drop_na() 

# get encode labels
labels9 <- get_labels(figure9.1$sp_area)
figure9.1$sp_area <- map_chr(.x = figure9$sp_area, .f = ~labels[.x])

# remove numbers from label
figure7 <- figure7 %>% 
  mutate(sa_policy_adap2 = str_remove(sa_policy_adap2,"[1-9].[1-9]. "))

# combine label
figure7$combined <- 
  paste(figure7$country_name, figure7$sa_policy_adap2, sep = " (")

figure7$combined <-
  paste(figure7$combined, ")", sep = "")

# plot the graph
figure7 %>% 
  mutate(combined = forcats::fct_reorder(.x = sa_app,
                                         .f = combined)) %>% 
  ggplot() +
  geom_col(aes(x = combined, y = sa_app),
           fill = "darkgoldenrod2",
           width = 0.4) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Figure 7. Top 10 cash transfer programs by actual coverage (% population)",
       y = "% of population covered by program",
       x = "") +
  theme(plot.title = element_text(face = "bold", size = 11, vjust = 1, hjust = 1.2),
        axis.title.x = element_text(size = 10))

ggsave(here("output/data_viz", "figure_9.png"))
```
